---
title: "L08 Coords"
subtitle: "Data Visualization (STAT 302)" 
author: "Shuo Han"

format:
  html:
    toc: true
    embed-resources: true
    code-fold: show
    link-external-newwindow: true
    fig-dpi: 300

execute:
  warning: false

from: markdown+emoji 
---

## Overview

The goal of this lab is to use an alternative coordinate system in `ggplot2` to build a plot. 

## Dataset

We'll be using the `mod_nba2014_15_advanced.txt` dataset --- add to your project's `/data` subdirectory. The `codebook_mod_nba2014_15_advanced.txt` provides a quick description of the variables in the dataset --- suggest adding it to the `/data` subdirectory as well. 

```{r}
#| label: load-pkgs-data

# load package(s)
library(tidyverse)

# load data
nba_data <- read_delim("data/mod_nba2014_15_advanced.txt", delim = "|") %>%
  janitor::clean_names()
```

## Exercise 

<!-- See the html instructions since they are complete with images and are the official instructions. -->

Using the `mod_nba2014_15.txt` dataset we will try to recreate/approximate the plot type featured in the [http://fivethirtyeight.com/](fivethirtyeight.com) article [Kawhi Leonard Is The Most Well-Rounded Elite Shooter Since Larry Bird](http://fivethirtyeight.com/features/kawhi-leonard-is-the-most-well-rounded-elite-shooter-since-larry-bird/) for any player of your choice for the 2014-2015 season. 

**Replicate the graphic below as closely as possible:**

### Start with data wrangling

When data wrangling we will need a helper function for creating the quartiles for players. Additionally, it will be useful to exclude players that played less than 10 games or played less than 5 minutes a game. That is, we only want to include "qualified" players.

```{r}
#| label: helper-function

# Returns quartile rank
quartile_rank <- function(x = 0:99) {
  # Set quartile
  quart_breaks <- c(
    -Inf,
    quantile(x,
      probs = c(.25, .5, .75),
      na.rm = TRUE
    ),
    Inf
  )
  cut(x = x, breaks = quart_breaks, labels = FALSE)
}
```

:::{.callout-important}
**Add comments to** the code below where indicated. The added comments should concisely describe what the following line(s) of code do in the data wrangling process  
:::

::: {.callout-tip icon=false}

## Solution

```{r}
#| label: data-wrangling

# Graphical data
nba_graph_dat <- nba_data %>% 
  # filter to get only players played in fewer than 10 games or fewer than 5 minutes per game
  filter(g >= 10, mp / g >= 5) %>% 
  # creates new columns in the data frame for each skill's quartile rank
  mutate(
    ts_quant = quartile_rank(ts_perc),
    trb_quant = quartile_rank(trb_perc),
    dbpm_quant = quartile_rank(dbpm),
    ast_quant = quartile_rank(ast_perc),
    usg_quant = quartile_rank(usg_perc)
  ) %>% 
  # select only the player column and the columns containing _quant
  select(player, contains("_quant")) %>% 
  # converts the data frame from wide to long format, pivoting all columns except for player, and creating a new column called skill to store the original column names, and a new column called value to store the values from the original columns
  pivot_longer(
    cols = -player, 
    names_to = "skill", 
    values_to = "value"
    ) %>% 
  # sort the data frame by player
  arrange(player)
```

:::

### Build the plot

*Hints:*

- Image width is 8 in, height is 10 in --- use code chunk yaml
- center the image output --- use code chunk yaml
- shrink size down for display to 50% width --- use code chunk yaml
- All added text, except title, is done with `annotate()`
- set `axis.text` to blank in theme
- size 3 and 5 could be useful

::: {.callout-tip icon=false}
Challenge plot is shown below, so not show the original one here.
:::

## Challenge

<!-- See the html instructions since they are complete with images and are the official instructions. -->

**Challenge is optional for all students, but we recommend trying them out!** 

Continue the work of trying to recreate the graphic from the article. There is a pre built theme from `ggthemes` that might be useful. Your final plot should look as follows:

::: {.callout-tip icon=false}

## Solution
```{r}
#| label: skills-chart
#| fig-width: 8
#| fig-height: 10
#| fig-align: center
#| out-width: "50%"

# identify player
player_name <- "LeBron James"

# plot data
plot_data <- nba_graph_dat %>%
  filter(player == player_name)

# skills chart
ggplot(plot_data, aes(x = skill, y = value)) +
  geom_col(width = 1, fill = "#F07786") +
  geom_segment(x = 0.5:4.5, y = 0, xend = 0.5:4.5, yend = 4) +
  geom_hline(yintercept = 1:4, linetype = "dotted") +
  scale_x_discrete(
    name = NULL,
    limits = c("ts_quant", "usg_quant", "dbpm_quant", "trb_quant", "ast_quant"),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    name = NULL,
    expand = c(0, 0)
  ) +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid.major = element_blank(),
    plot.title = element_text(size = 24, hjust = 0.5)
  ) +
  # sector labels
  annotate(
    geom = "text",
    label = c("TRUE\nSHOOTING", "USAGE\nRATE", "DEFENSIVE\nBPM", "REBOUND\nRATE", "ASSIST\nRATE"),
    x = 1:5,
    y = 5,
    size = 5
  ) +
  # adding quartile labels
  annotate(
    geom = "text",
    label = c("1st-25th", "25th-50th", "50th-75th", "75th-99th"),
    x = 3,
    y = 0.7:3.7,
    size = 3
  ) +  
  labs(
    title = str_c(player_name, "\n", "(2015)")
  ) +
  coord_polar()
```

:::